The Ontological Modal Collapse \\ as a Collapse of the Square of Opposition

### 2014-08-28 Papers/Universalis/ModalCollapse.tex
Structure paper and write introduction

\documentclass{birkjour}
%
%
% THEOREM Environments (Examples)-----------------------------------------
%
 \newtheorem{thm}{Theorem}[section]
 \newtheorem{cor}[thm]{Corollary}
 \newtheorem{lem}[thm]{Lemma}
 \newtheorem{prop}[thm]{Proposition}
 \theoremstyle{definition}
 \newtheorem{defn}[thm]{Definition}
 \theoremstyle{remark}
 \newtheorem{rem}[thm]{Remark}
 \newtheorem*{ex}{Example}
 \numberwithin{equation}{section}

@@

\title[Modal Collapse]
 {The Ontological Modal Colapse \\ 
 as a Collapse of the Square of Opposition}

@@

\address{%
Viaduktstr. 42\\
P.O. Box 133\\
CH 4010 Basel\\
Switzerland}

@@

%----------Author 2
\author[Woltzenlogel-Paleo]{Bruno Woltzenlogel Paleo}
\address{ }
\email{bruno.wp@gmail.com}
%----------classification, keywords, date
\subjclass{
Primary 03A02;  % Philosophical aspects of logic and foundations
Secondary 68T02 % Artificial Intelligence
}

\keywords{Modal Logics, Higher-Order Modal Logics}

\date{August 21, 2014}
%----------additions
\dedicatory{ }
%%% ----------------------------------------------------------------------

\begin{abstract}
This work presents.
\end{abstract}

@@

\begin{defn}
This serves as environment for definitions. Note that the text
appears not in italics.
\end{defn}

\begin{equation}\label{testequation}
\text{This is a sample equation: } c^2=a^2+b^2
\end{equation}

\begin{thm}[Main Theorem]
In contrast to definitions, theorems appear typeset in italics as
it has become more or less standard in most textbooks and
monographs. Equations can be cited using the \verb+\eqref+ command which
automatically adds brackets: \verb+\eqref{testequation}+ results in \eqref{testequation}.
\end{thm}

\begin{proof}
A special environment is predefined: the \textit{proof} environment. Please use
\begin{verbatim}\begin{proof}\end{verbatim}
proof of the statement
\begin{verbatim}\end{proof}\end{verbatim}
for typesetting your proofs. The end-of-proof symbol $\Box$ will be added automatically.
\end{proof}

There are two known problems with the placement of the end-of-proof sign:

\begin{enumerate}
  \item if your proof ends with a\ \ s i n g l e\ \ displayed line, the end-of-proof sign would
be placed in the line below; if you want to avoid this, write your line in the form
\begin{verbatim}$$displayed math line \eqno\qedhere$$\end{verbatim}
which results in

\begin{proof}
$$displayed math line \eqno\qedhere$$
\end{proof}
\item if your proof ends with an aligned displayed environment, the command
\verb+\tag*{\qed}+ can be used to place the end-of-proof sign properly:
\begin{verbatim}
\begin{align*}
\alpha&=\beta+\gamma\\
&=\delta+\epsilon\tag*{\qed}
\end{align*}
\end{verbatim}
results in
\begin{align*}
\alpha&=\beta+\gamma\\
&=\delta+\epsilon\tag*{\qed}
\end{align*}
\end{enumerate}
Please try to avoid using the obsolete \verb+\eqnarray+ environment. This environment has several bugs
and has been replaced by the more flexible \AmS\ environments \verb+align, split, multline+.

\begin{rem}
Additional comments are being typeset without boldfaced entrance
word as they may be minor important.
\end{rem}

\begin{ex}
For some constructs, even no number is required.
\end{ex}

Displayed equations may be numbered like the following one:
\begin{equation}
\sqrt{1-\sin^2(x)}=|\cos(x)|.
\end{equation}

\subsection{Conclusions}
% ------------------------------------------------------------------------

@@

\begin{thebibliography}{1}
\bibitem{test} A. B. C. Test, \textit{On a Test.} J. of Testing
\textbf{88} (2000), 100--120.
\bibitem{latex} G. Gr\"atzer, \textit{Math into \LaTeX.} 3rd Edition,
Birkh\"auser, 2000.
\end{thebibliography}

>>>>

\documentclass{birkjour}

\usepackage{url}
\usepackage{commands}
%
%
% THEOREM Environments (Examples)-----------------------------------------
%
 \newtheorem{thm}{Theorem}[section]
 \newtheorem{cor}[thm]{Corollary}
 \newtheorem{lem}[thm]{Lemma}
 \newtheorem{prop}[thm]{Proposition}
 \theoremstyle{definition}
 \newtheorem{defn}[thm]{Definition}
 \theoremstyle{remark}
 \newtheorem{rem}[thm]{Remark}
 \newtheorem*{ex}{Example}
 \numberwithin{equation}{section}

\def\HOML{\entity{HOML}\xspace}
\def\HOL{\entity{HOL}\xspace}

@@

\title[Modal Collapse]
 {The Ontological Modal Collapse \\ 
 as a Collapse of the Square of Opposition}

@@

\address{%
Department of Mathematics and Computer Science\\
Arnimallee 7 \\
Room 115 \\
14195 Berlin \\
Germany
}

@@

%----------Author 2
\author[Weber]{\\ Leon Weber}
\address{ 
ToDo \\
ToDo \\
ToDo Berlin \\
Germany
}
\email{leon.weber@fu-berlin.de}

%----------Author 3
\author[Woltzenlogel-Paleo]{\\ Bruno Woltzenlogel Paleo}
\address{ 
Favoritenstra{\ss}e 9 \\
Room HA0402 \\
1040 Wien \\
Austria
}
\email{bruno.wp@gmail.com}

%----------classification, keywords, date
\subjclass{
Prim. 03A02;  % Philosophical aspects of logic and foundations
Sec. 68T02 % Artificial Intelligence
}

\keywords{Modal Logics, Higher-Order Logics, Ontological Argument}

\date{August 30, 2014}
%----------additions
\dedicatory{ }
%%% ----------------------------------------------------------------------

\begin{abstract}
The \emph{modal collapse} that afflicts G\"odel's modal ontological 
argument for God's existence is discussed from the perspective of the 
modal square of opposition. Furthermore, a computer-assisted verification 
of the claims that the emendations by Anderson and by Frode are immune to 
the modal collapse is presented.
\end{abstract}

@@

Attempts to prove the
existence (or non-existence) of God by means of abstract, ontological
arguments are an old tradition in western philosophy, with contributions by several prominent philosophers, including St. Anselm of
Canterbury, Descartes and Leibniz. Kurt G{\"o}del studied and further improved this argument, bringing it to a mathematically more precise form, as a chain of axioms, lemmas and theorems in a modal logic \cite{GoedelNotes,ScottNotes}, as in Fig. \ref{fig:scott}.

\begin{figure}[t]
\noindent \framebox[\columnwidth][r]{
\begin{minipage}{.94\columnwidth}\small
\begin{itemize}
\item[A1] Either a property or its negation is positive, but not
  both:
  $$\hol{\allq \phi [P(\neg \phi) \biimp \neg P(\phi)]}$$ 
\item[A2] A property necessarily implied by a
  positive property is positive:
  $$\hol{\allq \phi \allq \psi [(P(\phi) \wedge \nec \allq x [\phi(x)
  \imp \psi(x)]) \imp P(\psi)]}$$
\item[T1] Positive properties are possibly exemplified: 
  $$\hol{\allq \phi [P(\phi) \imp \pos \exq x \phi(x)]}$$ 
\item[D1] A \emph{God-like} being possesses all positive properties: 
  $$\hol{G(x) \biimp \forall \phi [P(\phi) \imp \phi(x)]}$$ 
\item[A3]  The property of being God-like is positive: 
  $$\hol{P(G)}$$
\item[C\phantom{1}] Possibly, a God-like being exists: $$\hol{\pos \exq x G(x)}$$
\item[A4]  Positive properties are necessarily positive: 
  $$\hol{\allq \phi [P(\phi) \imp \Box \; P(\phi)]}$$ 
\item[D2] An \emph{essence} of an individual is a property possessed by it and necessarily implying any of its properties: $$\hol{\ess{\phi}{x} \biimp \phi(x) \wedge \allq
  \psi (\psi(x) \imp \nec \allq y (\phi(y) \imp \psi(y)))}$$ 
\item[T2]  Being God-like is an essence of any
  God-like being: $$\hol{\allq x [G(x) \imp \ess{G}{x}]}$$
\item[D3] \emph{Necessary existence} of an individual is the necessary exemplification of all its essences: 
  $$\hol{\NE(x) \biimp \allq \phi [\ess{\phi}{x} \imp \nec
  \exq y \phi(y)]}$$
\item[A5] Necessary existence is a positive property: $$\hol{P(\NE)}$$ 
\item[T3] Necessarily, a God-like being exists: $$\hol{\nec \exq x G(x)}$$ 
\end{itemize}
\end{minipage}
} \vskip-.5em
\caption{Scott's version of G\"odel's ontological argument \cite{ScottNotes}.\label{fig:scott}} 
\end{figure}

% Ontological arguments, for or against the existence of God,
% illustrate well an essential aspect of metaphysics: some (necessary) facts
% for our existing world are deduced by purely a priori, analytical means from some
% abstract definitions and axioms. % Contingent truths are to be
% distinguished from necessary truths.

G\"{o}del defines God (see Fig.~\ref{fig1}) as a being who possesses all \emph{positive}
properties.  He does not extensively discuss what positive properties
are, but instead he states a few reasonable (but debatable) axioms
that they should satisfy.  
The overall idea of G{\"o}del's proof is in the tradition of Anselm's
argument, who defined God as some entity of which nothing greater can be
conceived. Anselm argued that existence in the actual world would
make such an assumed being even greater; hence, by definition God must
exist. G{\"o}del's ontological argument is clearly related to this
reasoning pattern. However, it also tries to fix some fundamental
weaknesses in Anselm's work. For example, G{\"o}del explicitly proves
that God's existence is possible, which has been a basic assumption for
Anselm. Because of this, Anselm's argument has been
criticized as incomplete by Leibniz. Leibniz claimed that the
assumption should be derivable from the definition of God as a perfect
being and from the notion of perfection.  G{\"o}del's proof 
addresses this critique, and it also addresses Kant's objection that existence should not be treated as a predicate.

Nevertheless, G{\"o}del's work
still leaves room for criticism. In particular, his axioms are so
strong that they entail a \emph{modal collapse} 
\cite{ToDo:sobel1987?,sobel2004logic}: 
everything that is the case is so necessarily. 
There has been an impressive body of recent and ongoing
work (cf.~\cite{sobel2004logic,Fitting,anderson90:_some_emend_of_goedel_ontol_proof,AndersonGettings,ToDo:FrodeAndHajek,ContemporaryBibliography} and the references therein)
proposing solutions for the modal collapse. 
The goal of this paper is to present a computer-assisted analysis of 
the solutions proposed by Anderson and by Frode. The technique enabling this analysis is the embedding of quantified modal logics into higher-order logics \cite{J23,B9,C36}, for which automated theorem provers exist \cite{ToDo:leo,sattalax,isabelle,coq}. This technique has already been succesfully employed in the verification and reconstruction of G\"odel's proof \cite{todo:all our previous papers}, and a detailed mathematical description of the technique is available in \cite{todo:ECAI}.

\section{The Modal Collapse}

\begin{figure}[t]
\noindent \framebox[\columnwidth][r]{
\begin{minipage}{.94\columnwidth}\small
\begin{itemize}
\item[MC] Everything that is the case is so necessarily:
  $$\hol{\allq \phi [\phi \imp \nec \phi]}$$ 
\item[MCC] Modalities collapse:
  $$\hol{\allq \phi [\phi \biimp \pos \phi \biimp \nec \phi]}$$
\end{itemize}
\end{minipage}
} \vskip-.5em
\caption{Modal Collapse \cite{todo:Sobel1987,SobelBook}.\label{fig:collapse}} 
\end{figure}

\clearpage

\section{Anderson's Emendation}

\begin{figure}[t]
\noindent \framebox[\columnwidth][r]{
\begin{minipage}{.94\columnwidth}\small
\begin{itemize}
\item[A:A1] If a property is positive, its negation is not positive:
  $$\hol{\allq \phi [P(\phi) \imp \neg P(\neg \phi)]}$$ 
\item[A2] A property necessarily implied by a
  positive property is positive:
  $$\hol{\allq \phi \allq \psi [(P(\phi) \wedge \nec \allq x [\phi(x)
  \imp \psi(x)]) \imp P(\psi)]}$$
\item[T1] Positive properties are possibly exemplified: 
  $$\hol{\allq \phi [P(\phi) \imp \pos \exq x \phi(x)]}$$ 
\item[A:D1] A \emph{God-like} being possesses all positive properties: 
  $$\hol{G_A(x) \biimp \forall \phi [P(\phi) \biimp \nec \phi(x)]}$$ 
\item[A3]  The property of being God-like is positive: 
  $$\hol{P(G)}$$
\item[C\phantom{1}] Possibly, a God-like being exists: $$\hol{\pos \exq x G(x)}$$
\item[A4]  Positive properties are necessarily positive: 
  $$\hol{\allq \phi [P(\phi) \imp \Box \; P(\phi)]}$$ 
\item[A:D2] An \emph{essence} of an individual is a property possessed by it and necessarily implying any of its properties: $$\hol{\essA{\phi}{x} \biimp \allq
  \psi [\nec \psi(x) \biimp \nec \allq y (\phi(y) \imp \psi(y))]}$$ 
\item[T2]  Being God-like is an essence of any
  God-like being: $$\hol{\allq x [G(x) \imp \ess{G}{x}]}$$
\item[D3] \emph{Necessary existence} of an individual is the necessary exemplification of all its essences: 
  $$\hol{\NE(x) \biimp \allq \phi [\ess{\phi}{x} \imp \nec
  \exq y \phi(y)]}$$
\item[A5] Necessary existence is a positive property: $$\hol{P(\NE)}$$ 
\item[T3] Necessarily, a God-like being exists: $$\hol{\nec \exq x G(x)}$$ 
\end{itemize}
\end{minipage}
} \vskip-.5em
\caption{Anderson's Emendation \cite{ToDo:Anderson}.\label{fig:anderson}} 
\end{figure}

\clearpage

\section{Frode's Alternative}

\begin{figure}[t]
\noindent \framebox[\columnwidth][r]{
\begin{minipage}{.94\columnwidth}\small
\begin{itemize}
\item[A1] Either a property or its negation is positive, but not
  both:
  $$\hol{\allq \phi [P(\neg \phi) \biimp \neg P(\phi)]}$$ 
\item[A2] A property necessarily implied by a
  positive property is positive:
  $$\hol{\allq \phi \allq \psi [(P(\phi) \wedge \nec \allq x [\phi(x)
  \imp \psi(x)]) \imp P(\psi)]}$$
\item[T1] Positive properties are possibly exemplified: 
  $$\hol{\allq \phi [P(\phi) \imp \pos \exq x \phi(x)]}$$ 
\item[D1] A \emph{God-like} being possesses all positive properties: 
  $$\hol{G(x) \biimp \forall \phi [P(\phi) \imp \phi(x)]}$$ 
\item[A3]  The property of being God-like is positive: 
  $$\hol{P(G)}$$
\item[C\phantom{1}] Possibly, a God-like being exists: $$\hol{\pos \exq x G(x)}$$
\item[A4]  Positive properties are necessarily positive: 
  $$\hol{\allq \phi [P(\phi) \imp \Box \; P(\phi)]}$$ 
\item[D2] An \emph{essence} of an individual is a property possessed by it and necessarily implying any of its properties: $$\hol{\ess{\phi}{x} \biimp \phi(x) \wedge \allq
  \psi (\psi(x) \imp \nec \allq y (\phi(y) \imp \psi(y)))}$$ 
\item[T2]  Being God-like is an essence of any
  God-like being: $$\hol{\allq x [G(x) \imp \ess{G}{x}]}$$
\item[D3] \emph{Necessary existence} of an individual is the necessary exemplification of all its essences: 
  $$\hol{\NE(x) \biimp \allq \phi [\ess{\phi}{x} \imp \nec
  \exq y \phi(y)]}$$
\item[A5] Necessary existence is a positive property: $$\hol{P(\NE)}$$ 
\item[T3] Necessarily, a God-like being exists: $$\hol{\nec \exq x G(x)}$$ 
\end{itemize}
\end{minipage}
} \vskip-.5em
\caption{Frode's Alternative \cite{ToDo:Frode}.\label{fig:frode}} 
\end{figure}

\clearpage

\section{Other Solutions}

ToDo: Fitting

\section{Conclusions}

Various slightly different versions of
axioms and definitions have been considered by G\"{o}del and by
several philosophers who commented on his proof
(cf.~\cite{sobel2004logic,anderson90:_some_emend_of_goedel_ontol_proof,AndersonGettings,Fitting,Adams,ContemporaryBibliography}).

In theoretical philosophy, formal logical confrontations with such
ontological arguments had been so far (mainly) limited to paper
and pen.  Up to now, the use of computers was prevented, because the
logics of the available theorem proving systems were not expressive
enough to formalize the abstract concepts adequately. G{\"o}del's proof
uses, for example, a complex higher-order modal logic (\HOML) to handle
concepts such as \emph{possibility} and \emph{necessity} and to support
quantification over individuals and properties.

ToDo: Leibniz calculemus, Rushby, Zalta \cite{oppenheimera11,rushby13}

% \begin{defn}
% This serves as environment for definitions. Note that the text
% appears not in italics.
% \end{defn}

% \begin{equation}\label{testequation}
% \text{This is a sample equation: } c^2=a^2+b^2
% \end{equation}

% \begin{thm}[Main Theorem]
% In contrast to definitions, theorems appear typeset in italics as
% it has become more or less standard in most textbooks and
% monographs. Equations can be cited using the \verb+\eqref+ command which
% automatically adds brackets: \verb+\eqref{testequation}+ results in \eqref{testequation}.
% \end{thm}

% \begin{proof}
% A special environment is predefined: the \textit{proof} environment. Please use
% \begin{verbatim}\begin{proof}\end{verbatim}
% proof of the statement
% \begin{verbatim}\end{proof}\end{verbatim}
% for typesetting your proofs. The end-of-proof symbol $\Box$ will be added automatically.
% \end{proof}

% There are two known problems with the placement of the end-of-proof sign:

% \begin{enumerate}
%   \item if your proof ends with a\ \ s i n g l e\ \ displayed line, the end-of-proof sign would
% be placed in the line below; if you want to avoid this, write your line in the form
% \begin{verbatim}$$displayed math line \eqno\qedhere$$\end{verbatim}
% which results in

% \begin{proof}
% $$displayed math line \eqno\qedhere$$
% \end{proof}
% \item if your proof ends with an aligned displayed environment, the command
% \verb+\tag*{\qed}+ can be used to place the end-of-proof sign properly:
% \begin{verbatim}
% \begin{align*}
% \alpha&=\beta+\gamma\\
% &=\delta+\epsilon\tag*{\qed}
% \end{align*}
% \end{verbatim}
% results in
% \begin{align*}
% \alpha&=\beta+\gamma\\
% &=\delta+\epsilon\tag*{\qed}
% \end{align*}
% \end{enumerate}
% Please try to avoid using the obsolete \verb+\eqnarray+ environment. This environment has several bugs
% and has been replaced by the more flexible \AmS\ environments \verb+align, split, multline+.

% \begin{rem}
% Additional comments are being typeset without boldfaced entrance
% word as they may be minor important.
% \end{rem}

% \begin{ex}
% For some constructs, even no number is required.
% \end{ex}

% Displayed equations may be numbered like the following one:
% \begin{equation}
% \sqrt{1-\sin^2(x)}=|\cos(x)|.
% \end{equation}

% ------------------------------------------------------------------------

@@

\begin{thebibliography}{1}

\bibitem{Adams}
R.M. Adams, `Introductory note to *1970', in {\em {Kurt G\"odel: Collected
  Works Vol. 3: Unpubl. Essays and Letters}}, Oxford Univ. Press, (1995).

\bibitem{AndersonGettings}
A.C. Anderson and M.~Gettings, `G\"odel ontological proof revisited', in {\em
  {G\"odel'96: Logical Foundations of Mathematics, Computer Science, and
  Physics: Lecture Notes in Logic 6}},  167--172, {Springer}, (1996).

\bibitem{anderson90:_some_emend_of_goedel_ontol_proof}
C.A. Anderson, `Some emendations of {G{\"o}del's} ontological proof', {\em
  Faith and Philosophy}, {\bf 7}(3), (1990).

\bibitem{Andrews:gmae72}
P.B. Andrews, `General models and extensionality', {\em Journal of Symbolic
  Logic}, {\bf 37}(2),  395--397, (1972).

\bibitem{andrewsSEP}
P.B. Andrews, `Church's type theory', in {\em The Stanford Encyclopedia of
  Philosophy}, ed., E.N. Zalta, spring 2014 edn., (2014).

\bibitem{C36}
C.~Benzm{\"u}ller, `{HOL} based universal reasoning', in {\em Handbook of the
  4th World Congress and School on Universal Logic}, ed., J.Y. Beziau~et al.,
  pp. 232--233, Rio de Janeiro, Brazil, (2013).

\bibitem{B5}
C.~Benzm{\"u}ller and D.~Miller, `Automation of higher-order logic', in {\em
  Handbook of the History of Logic, Volume 9 --- Logic and Computation},
  Elsevier, (2014).
\newblock Forthcoming; preliminary version available at
  {http://christoph-benzmueller.de/papers/B5.pdf}.

\bibitem{C34}
C.~Benzm{\"u}ller, J.~Otten, and Th. Raths, `Implementing and evaluating
  provers for first-order modal logics', in {\em Proc. of the 20th European
  Conference on Artificial Intelligence (ECAI)}, pp. 163--168, (2012).

\bibitem{B9}
C.~Benzm{\"u}ller and L.C. Paulson, `Exploring properties of normal multimodal
  logics in simple type theory with {LEO-II}', in {\em {Festschrift in Honor of
  {Peter B. Andrews} on His 70th Birthday}}, ed., C.~Benzm{\"u}ller~et al.,
  386--406, College Publications, (2008).

\bibitem{J23}
C.~Benzm{\"u}ller and L.C. Paulson, `Quantified multimodal logics in simple
  type theory', {\em Logica Universalis}, {\bf 7}(1),  7--20, (2013).

\bibitem{LEO-II}
C.~Benzm{\"u}ller, F.~Theiss, L.~Paulson, and A.~Fietzke, `{LEO-II} - a
  cooperative automatic theorem prover for higher-order logic', in {\em
  Proc.~of IJCAR 2008}, number 5195 in LNAI, pp. 162--170. Springer, (2008).

\bibitem{J30}
C.~Benzm{\"u}ller and B.~Woltzenlogel-Paleo, `Formalization, mechanization and
  automation of {G{\"o}del's} proof of {God's} existence', {\em
  arXiv:1308.4526}, (2013).

\bibitem{J28}
C.~Benzm\"uller and B.~Woltzenlogel-Paleo, `{G{\"o}del's God in Isabelle/HOL}',
  {\em Archive of Formal Proofs}, (2013).

\bibitem{W50}
C.~Benzm\"uller and B.~Woltzenlogel-Paleo, `G\"odel's {God} on the computer',
  in {\em Proceedings of the 10th International Workshop on the Implementation
  of Logics}, EPiC Series. EasyChair, (2013).
\newblock Invited abstract.

\bibitem{Coq}
Y.~Bertot and P.~Casteran, {\em {Interactive Theorem Proving and Program
  Development}}, Springer, 2004.

\bibitem{Nitpick}
J.C. Blanchette and T.~Nipkow, `Nitpick: A counterexample generator for
  higher-order logic based on a relational model finder', in {\em Proc. of ITP
  2010}, number 6172 in LNCS, pp. 131--146. Springer, (2010).

\bibitem{Satallax}
C.E. Brown, `Satallax: An automated higher-order prover', in {\em Proc. of
  IJCAR 2012}, number 7364 in LNAI, pp. 111 -- 117. Springer, (2012).

\bibitem{ContemporaryBibliography}
R.~Corazzon.
\newblock Contemporary~bibliography~on~ontological~arguments: {\scriptsize
  \url{http://www.ontology.co/biblio/ontological-proof-contemporary-biblio.htm}}.

\bibitem{Fitting}
M.~Fitting, {\em Types, Tableaux and G\"odel's God}, Kluwer, 2002.

\bibitem{fitting98}
M.~Fitting and R.L. Mendelsohn, {\em First-Order Modal Logic}, volume 277 of
  {\em Synthese Library}, Kluwer, 1998.

\bibitem{Gallin75}
D.~Gallin, {\em Intensional and Higher-Order Modal Logic}, North-Holland, 1975.

\bibitem{garbacz12:_prover_simpl_expal_away}
P.~Garbacz, `{PROVER9's} simplifications explained away', {\em Australasian
  Journal of Philosophy}, {\bf 90}(3),  585--592, (2012).

\bibitem{GoedelNotes}
K.~G\"odel, {\em Appx.A: Notes in Kurt G\"odel's Hand},  144--145.
\newblock In  \cite{sobel2004logic}, 2004.

\bibitem{Henkin50}
L.~Henkin, `Completeness in the theory of types', {\em Journal of Symbolic
  Logic}, {\bf 15}(2),  81--91, (1950).

\bibitem{homl}
R.~Muskens, `{Higher Order Modal Logic}', in {\em Handbook of Modal Logic},
  ed., P~Blackburn~et al.,  621--653, Elsevier, Dordrecht, (2006).

\bibitem{Isabelle}
T.~Nipkow, L.C. Paulson, and M.~Wenzel, {\em {Isabelle/HOL: A Proof Assistant
  for Higher-Order Logic}}, number 2283 in LNCS, Springer, 2002.

\bibitem{oppenheimera11}
P.E. Oppenheimer and E.N. Zalta, `A computationally-discovered simplification
  of the ontological argument', {\em Australasian Journal of Philosophy}, {\bf
  89}(2),  333--349, (2011).

\bibitem{rushby13}
J.~Rushby, `The ontological argument in {PVS}', in {\em Proc.~of CAV Workshop
  ``Fun With Formal Methods''}, St. Petersburg, Russia,, (2013).

\bibitem{Schulz:AICOM-2002}
S.~Schulz, `E -- a brainiac theorem prover', {\em {AI Communications}}, {\bf
  15}(2),  111--126, (2002).

\bibitem{ScottNotes}
D.~Scott, {\em Appx.B: Notes in Dana Scott's Hand},  145--146.
\newblock In  \cite{sobel2004logic}, 2004.

\bibitem{sobel2004logic}
J.H. Sobel, {\em Logic and Theism: Arguments for and Against Beliefs in God},
  Cambridge U. Press, 2004.

\bibitem{sutcliffe2009tptp}
G.~Sutcliffe, `The {TPTP} problem library and associated infrastructure', {\em
  Journal of Automated Reasoning}, {\bf 43}(4),  337--362, (2009).

\bibitem{J22}
G.~Sutcliffe and C.~Benzm{\"u}ller, `Automated reasoning in higher-order logic
  using the {TPTP THF} infrastructure.', {\em Journal of Formalized Reasoning},
  {\bf 3}(1),  1--27, (2010).

\bibitem{J29}
B.~Woltzenlogel-Paleo and C.~Benzm{\"u}ller, `Automated verification and
  reconstruction of {G\"odel's} proof of {God's} existence', {\em OCG J.},
  (2013).

\end{thebibliography}
